Constraints on Assembly Bias from Galaxy Clustering

### 2016-06-17 ms.tex

The results of the previous section suggest that halo mass is
insufficient to predict the galaxy content of a halo to the precision
mandated by an analysis of SDSS DR7 clustering measurements
(particularly the $\magr < -19$, $\magr < -20$, and $\magr < -20.5$
samples).  This should not be surprising, as the process of galaxy
formation and evolution is complex and there is little reason to
suspect that many environmental factors impact galaxy properties.
Our results further suggest that halo concentration (or some other
halo property correlated with halo concentration) impacts the galaxy
content of a halo, at least for galaxies of some luminosities. In
particular, the assembly bias inferred from the $\magr < -20$ and
$\magr < -20.5$ samples suggests that galaxies in roughly Milky
Way-sized halos tend to be more luminous if they reside in halos with
higher-than-average concentrations.

Within the context of the standard model of structure formation, this
can be interpreted in at least two ways that are not mutually
exclusive. Although our analysis cannot support either of these
scenarios unambiguously, they are reasonable starting points for
further consideration.  One possibility is simply that the growth and
structure of the host halo itself may drive this
luminosity-concentration dependence. For example, halos of higher
concentration at fixed mass have deeper potential wells (specifically,
higher escape velocities from the halo center).  Therefore, gravity
more strongly binds the stellar and gaseous contents of such halos
possibly driving more rapid star formation or making the galaxies less
susceptible to feedback mechanisms that suppress star
formation. Higher concentration halos have also formed earlier, on
average \citep[e.g.,][]{wechsler02}, which may also contribute to the
galaxy ending up more luminous. Halo assembly bias persists to large
separations ($r \gg \mathrm{Mpc}$) and this picture can therefore
accommodate galaxy assembly bias on such large scales.

@@

This ``splashback" hypothesis has further merit in that it has the
same sense as suggested by the data. Splashback halos have experienced
mass loss as a consequence of their tidal interactions with a larger,
nearby halo. As a consequence, splashback halos are expected to be
less massive than `regular' host halos that host similar galaxies
\citep{dalal_etal08}.  Moreover, high concentrations due to tidal
truncation are a hallmark of splashback halos
\citep{wang_etal07,dalal_etal08,sunayama_etal16}. This results in a
picture in which there is a population of halos, correlated on scales
$r \sim 1-3\, h^{-1}\mathrm{Mpc}$, that have concentrations that are
high for their mass and galaxies that are overluminous for their
mass. This is precisely the sense of assembly bias suggested by our
analysis. Note, though, that this splashback effect cannot explain
galaxy assembly bias on large, linear scales.

@@

We thank Zheng Zheng for his help in diagnosing the
differences between the present analysis and the SDSS DR7
analysis presented in \citet{zehavi_etal11}. We thank Manodeep 
Sinha for help installing and running the {\tt CorrFunc} package 
as well as comments on an earlier draft of this manuscript.
The authors gratefully acknowledge the Gauss Centre for Supercomputing
e.V. (www.gauss-centre.eu) and the Partnership for Advanced
Supercomputing in Europe (PRACE, www.prace-ri.eu) for funding the
MultiDark simulation project by providing computing time on the GCS
Supercomputer SuperMUC at Leibniz Supercomputing Centre (LRZ,
www.lrz.de). The Bolshoi simulations have been performed within the
Bolshoi project of the University of California High-Performance
AstroComputing Center (UC-HiPACC) and were run at the NASA Ames
Research Center. The work of ARZ was supported by the US National Science
Foundation via grant AST 1517563. FvdB is supported by the US National Science
Foundation through grant AST 1516962. APH is funded through the Yale Center 
for Astronomy \& Astrophysics Prize fellowship.

>>>>

The results of the previous section suggest that halo
mass is insufficient to predict the galaxy content of a halo to the
precision mandated by an analysis of SDSS DR7 clustering measurements
(particularly the $\magr < -19$, $\magr < -20$, and $\magr < -20.5$
samples).  This should not be surprising, as the process of galaxy
formation and evolution is complex and there is little reason to
suspect that all potentially-important factors are determined by a
single halo property. Our results further suggest that halo
concentration (or some other halo property correlated with halo
concentration) impacts the galaxy content of a halo, at least for
galaxies of some luminosities. In particular, the assembly bias
inferred from the $\magr < -20$ and $\magr < -20.5$ samples suggests
that galaxies in roughly Milky Way-sized halos tend to be more
luminous if they reside in halos with higher-than-average
concentrations.

Within the context of the standard model of structure formation, this
can be interpreted in at least two ways that are not mutually
exclusive. Although our analysis cannot support either of these
scenarios unambiguously, they are reasonable starting points for
further consideration.  One possibility is simply that the growth and
structure of the host halo itself may drive this
luminosity-concentration dependence. For example, halos of higher
concentration at fixed mass have deeper potential wells (specifically,
higher escape velocities from the halo center).  Therefore, gravity
more strongly binds the stellar and gaseous contents of such halos
possibly driving more rapid star formation or making the galaxies less
susceptible to feedback mechanisms that suppress star
formation. Higher concentration halos have also formed earlier, on
average \citep[e.g.,][]{wechsler02}. This suggests that any physics
that acts upon the baryonic content of the halo and its high-density
environment may have been operative for a longer time, resulting in a
larger, more luminous galaxy.  Concentration-dependent halo assembly
bias persists to very large separations \citep[$r > 10\, h^{-1}\mathrm{Mpc}$, e.g.,][]{wechsler06} and this picture can
therefore accommodate galaxy assembly bias on such large scales.

@@

This ``splashback" hypothesis has further merit in that it has the
same sense as suggested by the data. Splashback halos have experienced
mass loss as a consequence of their tidal interactions with a larger,
nearby halo. As a consequence, splashback halos are expected to be
less massive than `regular' host halos that host similar galaxies
\citep{dalal_etal08}.  Meanwhile, high concentrations due to tidal
truncation are a hallmark of splashback halos
\citep{wang_etal07,dalal_etal08,sunayama_etal16}.  The implication is
a population of halos, correlated on scales $r \sim 1-3\,
h^{-1}\mathrm{Mpc}$, that have concentrations that are high for their
mass and galaxies that are overluminous for their mass. This is
precisely the sense of assembly bias suggested by our analysis.  Note,
though, that this splashback effect cannot explain galaxy assembly
bias on large, linear scales.

Our formulation of the Decorated HOD is not designed to address 
the finer-grained causes for observed correlations. 
Rather, what we have provided in this work is a robust boundary condition 
that more detailed specific models, such as semi-analytical models 
and hydrodynamical simulations, should satisfy. 
However, we reiterate an important corollary to our findings. 
Although SDSS projected clustering measurements favor strong levels of assembly bias, 
the constraints on even our simple $\abias$ parameters are relatively weak. 
This implies that {\em it may not be possible to constrain a finer-grained physical 
picture using projected clustering at a level that is interesting.} 
In order to constrain models that are more complex than simple empirical models 
such as the Decorated HOD, it is necessary to leverage additional constraining power of 
observations beyond two-point clustering.

@@

We thank Zheng Zheng for his help in diagnosing the
differences between the present analysis and the SDSS DR7
analysis presented in \citet{zehavi_etal11}. We thank Manodeep 
Sinha for help installing and running the {\tt CorrFunc} package 
as well as comments on an earlier draft of this manuscript. ARZ and 
APH are grateful to Prince Rogers Nelson for inspiring our creativity. 
The authors gratefully acknowledge the Gauss Centre for Supercomputing
e.V. (www.gauss-centre.eu) and the Partnership for Advanced
Supercomputing in Europe (PRACE, www.prace-ri.eu) for funding the
MultiDark simulation project by providing computing time on the GCS
Supercomputer SuperMUC at Leibniz Supercomputing Centre (LRZ,
www.lrz.de). The Bolshoi simulations have been performed within the
Bolshoi project of the University of California High-Performance
AstroComputing Center (UC-HiPACC) and were run at the NASA Ames
Research Center. The work of ARZ was supported by the US National Science
Foundation via grant AST 1517563. FvdB is supported by the US National Science
Foundation through grant AST 1516962. APH is funded through the Yale Center 
for Astronomy \& Astrophysics Prize fellowship.
